---
title: "Machine learning - HF1"
author: "Viviane Schneider"
date: "7 de fevereiro de 2020"
data: "Sociology Group - HF1 Project"
output: 
  html_document: 
    highlight: zenburn
    keep_md: yes
    number_sections: yes
    theme: cerulean
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  cache = TRUE)
setwd("~/Machine learning")
```

# Study context

This is 

# Data Sources

You can 
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# The study goal


# Lybraries
```{r}
library(rattle)
library(caret)
library(rpart)
library(rpart.plot)
library(corrplot)
library(randomForest)
library(RColorBrewer)

```


# Get and read data

setwd("~/GitHub/Practical-Machine-Learning-Johns-Hopkins-Bloomberg-School-of-Public-Health-Coursera/Project")


```{r read in github}



trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile = trainFile, method = "curl")
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile = testFile, method = "curl")
}
rm(trainUrl)
rm(testUrl)

##Reading

trainRaw <- read.csv(trainFile)
testRaw <- read.csv(testFile)
dim(trainRaw)


```

# Cleaning

```{r}
##  clean the Near Zero Variance Variables.
NZV <- nearZeroVar(trainRaw, saveMetrics = TRUE)
head(NZV, 20)



training01 <- trainRaw[, !NZV$nzv]
testing01 <- testRaw[, !NZV$nzv]
dim(training01)


rm(trainRaw)
rm(testRaw)
rm(NZV)

## Removing some columns of the dataset that do not contribute much to the accelerometer measurements.

regex <- grepl("^X|timestamp|user_name", names(training01))
training <- training01[, !regex]
testing <- testing01[, !regex]
rm(regex)
rm(training01)
rm(testing01)
dim(training)

##Removing columns that contain NA's.

cond <- (colSums(is.na(training)) == 0)
training <- training[, cond]
testing <- testing[, cond]
rm(cond)


```

# Correlation

```{r}
corrplot(cor(training[, -length(names(training))]), method = "color", tl.cex = 0.5)

```



# Partitioning Training Set

```{r}
set.seed(56789) # For reproducibile purpose
inTrain <- createDataPartition(training$classe, p = 0.70, list = FALSE)
validation <- training[-inTrain, ]
training <- training[inTrain, ]
rm(inTrain)

```


# Data modeling

## Decision Tree

fit a predictive model for activity recognition using Decision Tree algorithm.

```{r}
modelTree <- rpart(classe ~ ., data = training, method = "class")
prp(modelTree)

```


estimate the performance of the model on the validation data set.

```{r}
predictTree <- predict(modelTree, validation, type = "class")
confusionMatrix(validation$classe, predictTree)

accuracy <- postResample(predictTree, validation$classe)
ose <- 1 - as.numeric(confusionMatrix(validation$classe, predictTree)$overall[1])
rm(predictTree)
rm(modelTree)


```

## Random Forest

fit a predictive model for activity recognition using Random Forest algorithm because it automatically selects important variables and is robust to correlated covariates & outliers in general.
We will use 5-fold cross validation when applying the algorithm.

```{r}
modelRF <- train(classe ~ ., data = training, method = "rf", trControl = trainControl(method = "cv", 5), ntree = 250)
modelRF

```

 estimate the performance of the model on the validation data set.
 
```{r}
predictRF <- predict(modelRF, validation)
confusionMatrix(validation$classe, predictRF)


accuracy <- postResample(predictRF, validation$classe)
ose <- 1 - as.numeric(confusionMatrix(validation$classe, predictRF)$overall[1])
rm(predictRF)

```
 
# Predicting The Manner of Exercise for Test Data Set

apply the Random Forest model to the original testing data set downloaded from the data source. We remove the problem_id column first.

```{r}
rm(accuracy)
rm(ose)
predict(modelRF, testing[, -length(names(testing))])

```

# Generating Files to submit as answers for the Assignment

Function to generate files with predictions to submit for assignment.

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("./Assignment_Solutions/problem_id_",i,".txt")
    write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, col.names = FALSE)
  }
}
```

 Generating the Files.
 

pml_write_files(predict(modelRF, testing[, -length(names(testing))]))
rm(modelRF)
rm(training)
rm(testing)
rm(validation)
rm(pml_write_files)

```
 
